Automated Formal Verification of Stand-alone Solar Photovoltaic Systems

### 2019-07-16 Tex/elsevier/trindade_solar_energy_rev1_unmarked.tex
solar energy revised text and reply letter

\begin{abstract}
\textcolor{red}{With declining costs and increasing performance, the deployment of renewable energy systems is growing faster. In $2017$, for the first time, the number of people without access to electricity dropped down below $1$ billion, but trends on energy access likewise fall short of global goals. Particular attention is given to stand-alone solar photovoltaic (PV) systems in rural areas or where grid extension is unfeasible. Tools to evaluate electrification projects are available, but they are based on simulations that do not cover all aspects of the design-space. Automated verification using model checking has proven to be an effective technique to validate complex (state transition) systems. This paper marks the first application of software model checking to formally verify the design of a stand-alone solar photovoltaic system, including solar panel, charge controller, battery, inverter, and electric load. Our main focus is on the solar PV project validation to be carried out just after the system sizing, i.e., prior to buying equipment and deploying it in the field, as a safe approach to ensure the intended behavior. Case studies, ranging from $253$ W to $814$ W, were used to evaluate this proposed approach and to compare that with specialized simulation tool. Different verification tools were evaluated to compare performance and soundness among automated verifiers. The experimental results reported by our automated verification method and by the simulation tool were compared with data collected from dwellers of the deployed cases, thereby showing the effectiveness of our approach, where specific conditions that lead to failures in a solar PV system are only detailed by the automated verification method.}
\end{abstract}

@@

\textcolor{red}{Note that in this study, our focus is not on a novel mathematical modelling of PV systems. Instead, our novelty relies on an effective approach to perform validation of PV systems using software model checking, in a process that is intended to be done just after the system sizing, prior to buying equipment and deploying it in the field, as a safe approach to ensure the intended behavior.}

\textcolor{red}{Here we compare model checking with simulation tools. However how they work is completely different. On one hand, simulation depends on the choice of input variables (and their values) in order to obtain the output. On the other hand, model checking performs an exhaustive search with the goal of proving or violating a given property. In the latter case, if the property is violated (an unexpected behavior of the system), then the model checker presents the input that causes such a violation. Therefore, we can prove the absence of system flaws using model checking.}

@@

As highlights, only HOMER and Hybrid2 perform off-grid system with battery backup analysis. \textcolor{red}{Hybrid2 is not supported anymore, being in disuse since that not work on Windows platforms later than Windows XP~\citep{UMASS, Pinho}.} %all the tools perform solar photovoltaic analysis; 

@@

A stand-alone PV system is illustrated in Fig.\ref{fig:blockdiagram}. The PV generator is a semiconductor device that can convert solar energy into DC electricity\textcolor{red}{, with high dependence on two weather variables from the site, where the system is deployed: solar irradiance $G$ and temperature $T$}. 
For night hours or rainy days, power stored in batteries can be used and it implies the presence of a charge controller~\citep{Hansen}. The PV arrays produce DC and therefore when the PV system contains an AC load, a DC/AC conversion is required (inverter). The AC load dictates the behavior of AC electrical load from the house that will be fed by the system. \textcolor{red}{In this pictured modular structure, every element produces/consumes current $I$ and voltage $V$, as illustrated by their physical magnitudes, where $pv$ means photovoltaic, $bat$ is battery, $dc$ is direct, and $ac$ is alternating signals.}

@@

\caption{\textcolor{red}{Block diagram for a typical stand-alone PV system, adapted from~\citep{Hansen}}.}

@@

%The performance of PV systems %, %that is, the output current/voltage curve ($I-V$ curve), 
%is usually studied using an equivalent circuit model~\cite{Yatimi}, which consists of a current source with one or two diodes connected in parallel, and up to two resistors, one connected in parallel and the other one in series, to take into account energy losses in this model~\cite{Cubas}. 
%Based on these electronic components, four basic configurations are normally used when studying photovoltaic systems, as shown by Fig. \ref{fig:equivckt}. 
\begin{figure}[h]
\includegraphics[width=0.40\textwidth]{equivckt1D}
\centering
\caption{\textcolor{red}{1-diode equivalent PV cell/panel circuit model, adapted from~\citep{Cubas}.}}
%\caption{Four different equivalent circuit models: (a) 1-diode; (b) 1-diode/1-resistor; (c) 1-diode/2-resistor; (d) 2-diode/2-resistor (Source: \cite{Cubas})}
\label{fig:equivckt}
\end{figure}

@@

\textcolor{red}{Note that the steps described here are carried out shortly after the design of the solar PV system, as soon as the equipment and its specifications are defined, i.e., before buying and deploying them, as a safe approach to ensure that the project will succeed.}

@@

In Step 1, the PV input data \textcolor{red}{(e.g., load power demand and load energy consumption)} 
and the formulae to check the sizing project, the mathematical model, the limits of the weather non-deterministic variables, are all written as an ANSI-C code~\citep{ANSI2018}. In Step 2, the sizing check of the PV system takes place: it will indicate if there is an error of sizing before to perform the automated verification of the system. This stage ensures that the system meets the standard project steps related to critical period method of sizing~\citep{Pinho}. 
%
In Step 3, weather variables (e.g., solar irradiance and ambient temperature) will be systematically explored by our verification engine based on maximum and minimum values from the site, where the PV system will be deployed. 
%As a consequence, all the formulae of the employed mathematical models will also be updated. 
In addition, depending on one of the desired properties of the system such as battery autonomy, energy availability, or even system power supply, our verification engine is able to indicate a failure if those properties are not met; in this particular case, it provides a diagnostic counterexample that shows in which conditions the property violation occurred. 
%; as the  state of charge of the batteries, load demand of power and the load consumption of energy if defined by the code
% (as reliability, performance, or safety)

@@

\textcolor{red}{In our study, we estimated load curves based on survey and visiting performed in June of 2017, and the the estimated load curve was converted as vectors with 24 positions, one to each house of the day (in Watts):}
%
\begin{itemize} 
\item \textcolor{red}{House 1: [118, 118, 118, 46, 46, 46, 95, 95, 170, 170, 296, 242, 242, 95, 95, 95, 95, 95, 342, 288, 288, 288, 288, 118];} 
\item \textcolor{red}{House 2: [136, 136, 136, 136, 136, 136, 67, 67, 184, 184, 184, 184, 184, 67, 67, 67, 67, 67, 253, 253, 253, 253, 253, 136];} 
\item \textcolor{red}{House 3: [113, 113, 113, 113, 113, 113, 67, 67, 217, 97, 97, 97, 97, 97, 97, 97, 97, 97, 263, 113, 113, 113, 113, 113];}
\item \textcolor{red}{House 4: [207, 207, 207, 135, 135, 135, 66, 66, 161, 161, 233, 253, 248, 66, 66, 66, 66, 66, 302, 317, 322, 302, 302, 207];} 
\item \textcolor{red}{House 5: [45, 16, 16, 16, 16, 16, 0, 0, 0, 72, 72, 222, 150, 150, 0, 0, 72, 72, 814, 814, 814, 742, 742, 16].}
\end{itemize}

@@

\section{\textcolor{red}{Verification and Simulation Results}} 

@@

We have performed five case studies to evaluate the proposed approach as described in Table~\ref{tab2}. \textcolor{red}{Power peak, power surge, and energy consumption were estimated based on visiting and survey applied to each house in June 2017 (before the electrification).} %Furthermore, three start-of-art verification tools, as described in Section~\ref{sec:AutomatedVerification} (ESBMC, CBMC, and CPAchecker), and HOMER Pro simulation tool were used to compare the approach effectiveness and efficiency.

\begin{table}
\caption{Case studies: stand-alone solar PV systems.}\label{tab2}
\begin{scriptsize}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\hline
Item & House 1 & House 2 & House 3 & House 4 & House 5\\
\hline
\hline
PV Panels &  \multicolumn{4}{|c|}{3$\times$325 W: (3S)} & 4$\times$325 W: (2S-2P) \\
\hline
Batteries & \multicolumn{4}{|c|}{\makecell{4$\times$220 Ah: (2S-2P)\\ autonomy: 48 h}} & \makecell{4$\times$120 Ah: (4S)\\ autonomy: 6 h}\\
\hline
Charge Controller & \multicolumn{5}{|c|}{With MPPT of 150 V/35 A}\\
\hline
Inverter & \multicolumn{4}{|c|}{700 W, surge: 1,600 W} & 1,200 W, surge: 1,600 W\\
\hline
\textcolor{red}{Power peak (W)}& 342  & 253  & 263 & 322 & 814 \\
\hline
\textcolor{red}{Power surge (W)}& 342 & 722 & 732 & 896 & 980\\
\hline
\makecell{\textcolor{red}{Consumption}\\ \textcolor{red}{(kWh/day)}}& \textcolor{red}{3.9} & \textcolor{red}{3.6} & \textcolor{red}{2.5} & \textcolor{red}{4.3} & \textcolor{red}{4.88} \\
\hline
GPS Coordinates & \multicolumn{4}{|c|}{\makecell{2$^{o}$44'50.0"S 60$^{o}$25'47.8"W\\}} & \makecell{3$^{o}$4'20.208"S \\60$^{o}$0'30.168"W}\\
\hline
Details & \multicolumn{4}{|c|}{\makecell{Riverside indigenous community\\Rural Area of Manaus - Brazil}} & \makecell{Urban house \\Manaus-Amazonas-Brazil}\\
\hline
\hline
\end{tabular}
\textcolor{red}{Legend: (S): Series; (P): Parallel.}
\end{scriptsize}
\end{table}
%------------------------------------------
\subsection{Objectives and Setup}
\label{sec:setup}
%------------------------------------------
\textcolor{red}{We aim} to answer two research questions:
%
\begin{enumerate}
\item[RQ1] \textbf{(soundness)} Does our automated verification approach provide correct results?
\item[RQ2] \textbf{(performance)} How do the verifiers compare to each other and to a simulation commercial tool?
\end{enumerate}

@@

\textcolor{red}{Note that the results presented here depend on the computer's processor and memory, the version of each software verifier (i.e., CBMC, ESBMC, CPAchecker), the parameters passed by the command-line, and the implemented model (software created by authors) to be solved. Additionally, any change from HOMER Pro to another simulation tool can also influence the measured time to obtain results.}

@@

%At the terminal, the command used to perform the verification is:
%\$ esbmc filename.c -\phantom{}-no-bounds-check -\phantom{}-no-pointer-check -\phantom{}-no-div-by-zero-check -\phantom{}-unwind 300 -\phantom{}-smt-during-symex -\phantom{}-smt-symex-guard --z3
%
%Where:
%
%\begin{itemize}
%\item The first three parameters, after the filename, are related to options that are usual to find bug in software, %for example, like bound check to arrays, pointer check, and division by zero, 
%but unnecessary to check at this kind of problem (if not removed, there is lost of performance during the automated verification);
%\item The parameter $unmind$ tells to ESBMC the limit to unroll the loops. This number was optimized (empirically) in order to reduce the running time and avoid to unwind unnecessarily the loops;
%\item The two parameters with $symex$ tell to ESBMC to perform an incremental SMT solving. There are other options, but this parameter is necessary because the complexity of the algorithms. The incremental SMT solving uses few RAM memory, compared with other SMT solving. 
%During empirical tests of the algorithms, the incremental solving was the only one who do not demanded 100\% the RAM memory. The use of swap-memory, i.e., the use of hard disk, reduces the performance and must be avoided;
%\item And the last parameter says to the tool that the Z3 SMT solver will be used.
%\end{itemize}
%
%\subsection{Experimental results}
%\label{sec:results}
%---------------------------------------------------
\subsection{Threats to Validity}
%---------------------------------------------------
We have reported a favorable assessment of the proposed method. % over a diverse set of real-world benchmarks. 
Nevertheless, we have also identified \textcolor{red}{five} threats to the validity of our results that can further be assessed.

@@

\textcolor{red}{
\textit{Simulation Tool:} Only HOMER Pro was used. The inclusion of other specialized simulation tool or even a general simulation tool that uses the same mathematical model adopted by the automated verification could change the comparative.}

\textcolor{red}{
\textit{Temperature and Solar Irradiance Data}: Information related to temperature and solar irradiance of each case study, independently of using simulation or formal verification, come from databases available online~\citep{Temperature, Irradiance}. However, considering that riverside communities do not have weather stations, the data used in our study come from the closest municipality (Manaus in all case studies), where stations collect regularly those data. Therefore, the most accurate should be the use of weather stations in each location.}

@@

\textcolor{red}{Our main focus was on the method to validate PV system projects, based on software model checking, which can cover the design-space better than simulation tools. 
We have considered five case studies, ranging from \textcolor{red}{$253$ W to $814$ W}; three state-of-art verification engines were considered (ESBMC, CBMC, and CPAchecker); and one specialized simulation tool (HOMER Pro). The experimental results from automated verification and simulation tools were compared with information collected from dwelles of the sites, where the real PV systems were deployed.} Although the verification method proposed takes longer than simulation methods, it is able to present details that lead to failures in a PV system that is not a feature presented in commercial simulation tool. In particular, the proposed method was successful in finding sizing errors and indicating in details. Related to the verification engines comparative, the ESBMC with the Z3 solver executed in the incremental configuration presented the better performance (around four times faster than CBMC), used less RAM memory (less than 2 GB when compared to 9.2 GB of CBMC and 19.2 GB of CPAchecker), and all the results were sound because the PV owners and the monitoring system validated the possible flaws that the system could be presenting in the field.

@@

As future work, we will \textcolor{red} {expand the number of case studies (preferentially in different countries), develop the code of a general purpose simulation tool to include in the comparative, and also consider} other types of renewable energy and even hybrid ones to allow our method to verify typical rural electrification.

>>>>

\begin{abstract}
With declining costs and increasing performance, the deployment of renewable energy systems is growing faster. In $2017$, for the first time, the number of people without access to electricity dropped down below $1$ billion, but trends on energy access likewise fall short of global goals. Particular attention is given to stand-alone solar photovoltaic systems in rural areas or where grid extension is unfeasible. Tools to evaluate electrification projects are available, but they are based on simulations that do not cover all aspects of the design-space. Automated verification using model checking has proven to be an effective technique to validate complex (state transition) systems. This paper marks the first application of software model checking to formally verify the design of a stand-alone solar photovoltaic system, including solar panel, charge controller, battery, inverter, and electric load. Our main focus is on the project validation to be carried out just after the system sizing, i.e., prior to buying equipment and deployment, as a safe approach to ensure the intended behavior. Five case studies were used to evaluate this proposed approach and to compare that with specialized simulation tool. Different verification tools were evaluated to compare performance and soundness among automated verifiers. The experimental results reported by our automated verification method and by the simulation tool were compared with data collected from dwellers of the deployed cases, thereby showing the effectiveness of our approach, where specific conditions that lead to failures in a solar photovoltaic system are only detailed by the automated verification method.
%With declining costs and increasing performance, the deployment of renewable energy systems is growing faster. In $2017$, for the first time, the number of people without access to electricity dropped down below $1$ billion, but trends on energy access likewise fall short of global goals. Particular attention is given to stand-alone solar photovoltaic (PV) systems in rural areas or where grid extension is unfeasible. Tools to evaluate electrification projects are available, but they are based on simulations that do not cover all aspects of the design-space. Automated verification using model checking has proven to be an effective technique to validate complex (state transition) systems. This paper marks the first application of software model checking to formally verify the design of a stand-alone solar photovoltaic system, including solar panel, charge controller, battery, inverter, and electric load. Our main focus is on the solar PV project validation to be carried out just after the system sizing, i.e., prior to buying equipment and deploying it in the field, as a safe approach to ensure the intended behavior. Case studies, ranging from $253$ W to $814$ W, were used to evaluate this proposed approach and to compare that with specialized simulation tool. Different verification tools were evaluated to compare performance and soundness among automated verifiers. The experimental results reported by our automated verification method and by the simulation tool were compared with data collected from dwellers of the deployed cases, thereby showing the effectiveness of our approach, where specific conditions that lead to failures in a solar PV system are only detailed by the automated verification method.}
\end{abstract}

@@

Note that in this study, our focus is not on a novel mathematical modelling of PV systems. Instead, our novelty relies on an effective approach to perform validation of PV systems using software model checking, in a process that is intended to be done just after the system sizing, prior to buying equipment and deploying it in the field, as a safe approach to ensure the intended behavior.

Here we compare model checking with simulation tools. However how they work is completely different. On one hand, simulation depends on the choice of input variables (and their values) in order to obtain the output. On the other hand, model checking performs an exhaustive search with the goal of proving or violating a given property. In the latter case, if the property is violated (an unexpected behavior of the system), then the model checker presents the input that causes such a violation. Therefore, we can prove the absence of system flaws using model checking.

@@

As highlights, only HOMER and Hybrid2 perform off-grid system with battery backup analysis.  Hybrid2 is not supported anymore, being in disuse since that not work on Windows platforms later than Windows XP~\citep{UMASS, Pinho}. %all the tools perform solar photovoltaic analysis; 

@@

A stand-alone PV system is illustrated in Fig.\ref{fig:blockdiagram}. The PV generator is a semiconductor device that can convert solar energy into DC electricity, with high dependence on two weather variables from the site, where the system is deployed: solar irradiance $G$ and temperature $T$. 
For night hours or rainy days, power stored in batteries can be used and it implies the presence of a charge controller~\citep{Hansen}. The PV arrays produce DC and therefore when the PV system contains an AC load, a DC/AC conversion is required (inverter). The AC load dictates the behavior of AC electrical load from the house that will be fed by the system. In this pictured modular structure, every element produces/consumes current $I$ and voltage $V$, as illustrated by their physical magnitudes, where $pv$ means photovoltaic, $bat$ is battery, $dc$ is direct, and $ac$ is alternating signals.

@@

\caption{Block diagram for a typical stand-alone PV system, adapted from~\citep{Hansen}.}

@@

%The performance of PV systems %, %that is, the output current/voltage curve ($I-V$ curve), 
%is usually studied using an equivalent circuit model~\cite{Yatimi}, which consists of a current source with one or two diodes connected in parallel, and up to two resistors, one connected in parallel and the other one in series, to take into account energy losses in this model~\cite{Cubas}. 
%Based on these electronic components, four basic configurations are normally used when studying photovoltaic systems, as shown by Fig. \ref{fig:equivckt}. 
\begin{figure}[h]
\includegraphics[width=0.40\textwidth]{equivckt1D}
\centering
\caption{1-diode equivalent PV cell/panel circuit model, adapted from~\citep{Cubas}.}
%\caption{Four different equivalent circuit models: (a) 1-diode; (b) 1-diode/1-resistor; (c) 1-diode/2-resistor; (d) 2-diode/2-resistor (Source: \cite{Cubas})}
\label{fig:equivckt}
\end{figure}

@@

Note that the steps described here are carried out shortly after the design of the solar PV system, as soon as the equipment and its specifications are defined, i.e., before buying and deploying them, as a safe approach to ensure that the project will succeed.

@@

In Step 1, the PV input data (e.g., load power demand and load energy consumption) 
and the formulae to check the sizing project, the mathematical model, the limits of the weather non-deterministic variables, are all written as an ANSI-C code~\citep{ANSI2018}. In Step 2, the sizing check of the PV system takes place: it will indicate if there is an error of sizing before to perform the automated verification of the system. This stage ensures that the system meets the standard project steps related to critical period method of sizing~\citep{Pinho}. 
%
In Step 3, weather variables (e.g., solar irradiance and ambient temperature) will be systematically explored by our verification engine based on maximum and minimum values from the site, where the PV system will be deployed. 
%As a consequence, all the formulae of the employed mathematical models will also be updated. 
In addition, depending on one of the desired properties of the system such as battery autonomy, energy availability, or even system power supply, our verification engine is able to indicate a failure if those properties are not met; in this particular case, it provides a diagnostic counterexample that shows in which conditions the property violation occurred. 
%; as the  state of charge of the batteries, load demand of power and the load consumption of energy if defined by the code
% (as reliability, performance, or safety)

@@

In our study, we estimated load curves based on survey and visiting performed in June of 2017, and the the estimated load curve was converted as vectors with 24 positions, one to each house of the day (in Watts):
%
\begin{itemize} 
\item House 1: [118, 118, 118, 46, 46, 46, 95, 95, 170, 170, 296, 242, 242, 95, 95, 95, 95, 95, 342, 288, 288, 288, 288, 118];
\item House 2: [136, 136, 136, 136, 136, 136, 67, 67, 184, 184, 184, 184, 184, 67, 67, 67, 67, 67, 253, 253, 253, 253, 253, 136]; 
\item House 3: [113, 113, 113, 113, 113, 113, 67, 67, 217, 97, 97, 97, 97, 97, 97, 97, 97, 97, 263, 113, 113, 113, 113, 113];
\item House 4: [207, 207, 207, 135, 135, 135, 66, 66, 161, 161, 233, 253, 248, 66, 66, 66, 66, 66, 302, 317, 322, 302, 302, 207]; 
\item House 5: [45, 16, 16, 16, 16, 16, 0, 0, 0, 72, 72, 222, 150, 150, 0, 0, 72, 72, 814, 814, 814, 742, 742, 16].
\end{itemize}

@@

\section{Verification and Simulation Results} 

@@

We have performed five case studies to evaluate the proposed approach as described in Table~\ref{tab2}. Power peak, power surge, and energy consumption were estimated based on visiting and survey applied to each house in June 2017 (before the electrification). %Furthermore, three start-of-art verification tools, as described in Section~\ref{sec:AutomatedVerification} (ESBMC, CBMC, and CPAchecker), and HOMER Pro simulation tool were used to compare the approach effectiveness and efficiency.

\begin{table}
\caption{Case studies: stand-alone solar PV systems.}\label{tab2}
\begin{scriptsize}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\hline
Item & House 1 & House 2 & House 3 & House 4 & House 5\\
\hline
\hline
PV Panels &  \multicolumn{4}{|c|}{3$\times$325 W: (3S)} & 4$\times$325 W: (2S-2P) \\
\hline
Batteries & \multicolumn{4}{|c|}{\makecell{4$\times$220 Ah: (2S-2P)\\ autonomy: 48 h}} & \makecell{4$\times$120 Ah: (4S)\\ autonomy: 6 h}\\
\hline
Charge Controller & \multicolumn{5}{|c|}{With MPPT of 150 V/35 A}\\
\hline
Inverter & \multicolumn{4}{|c|}{700 W, surge: 1,600 W} & 1,200 W, surge: 1,600 W\\
\hline
Power peak (W)& 342  & 253  & 263 & 322 & 814 \\
\hline
Power surge (W)& 342 & 722 & 732 & 896 & 980\\
\hline
\makecell{Consumption\\(kWh/day)}& 3.9 & 3.6 & 2.5 & 4.3 & 4.88 \\
\hline
GPS Coordinates & \multicolumn{4}{|c|}{\makecell{2$^{o}$44'50.0"S 60$^{o}$25'47.8"W\\}} & \makecell{3$^{o}$4'20.208"S \\60$^{o}$0'30.168"W}\\
\hline
Details & \multicolumn{4}{|c|}{\makecell{Riverside indigenous community\\Rural Area of Manaus - Brazil}} & \makecell{Urban house \\Manaus-Amazonas-Brazil}\\
\hline
\hline
\end{tabular}
Legend: (S): Series; (P): Parallel.
\end{scriptsize}
\end{table}
%------------------------------------------
\subsection{Objectives and Setup}
\label{sec:setup}
%------------------------------------------
We aim to answer two research questions:
%
\begin{enumerate}
\item[RQ1] \textbf{(soundness)} Does our automated verification approach provide correct results?
\item[RQ2] \textbf{(performance)} How do the verifiers compare to each other and to a simulation commercial tool?
\end{enumerate}

@@

Note that the results presented here depend on the computer's processor and memory, the version of each software verifier (i.e., CBMC, ESBMC, CPAchecker), the parameters passed by the command-line, and the implemented model (software created by authors) to be solved. Additionally, any change from HOMER Pro to another simulation tool can also influence the measured time to obtain results.

@@

%At the terminal, the command used to perform the verification is:
%\$ esbmc filename.c -\phantom{}-no-bounds-check -\phantom{}-no-pointer-check -\phantom{}-no-div-by-zero-check -\phantom{}-unwind 300 -\phantom{}-smt-during-symex -\phantom{}-smt-symex-guard --z3
%
%Where:
%
%\begin{itemize}
%\item The first three parameters, after the filename, are related to options that are usual to find bug in software, %for example, like bound check to arrays, pointer check, and division by zero, 
%but unnecessary to check at this kind of problem (if not removed, there is lost of performance during the automated verification);
%\item The parameter $unmind$ tells to ESBMC the limit to unroll the loops. This number was optimized (empirically) in order to reduce the running time and avoid to unwind unnecessarily the loops;
%\item The two parameters with $symex$ tell to ESBMC to perform an incremental SMT solving. There are other options, but this parameter is necessary because the complexity of the algorithms. The incremental SMT solving uses few RAM memory, compared with other SMT solving. 
%During empirical tests of the algorithms, the incremental solving was the only one who do not demanded 100\% the RAM memory. The use of swap-memory, i.e., the use of hard disk, reduces the performance and must be avoided;
%\item And the last parameter says to the tool that the Z3 SMT solver will be used.
%\end{itemize}
%
%\subsection{Experimental results}
%\label{sec:results}
%---------------------------------------------------
\subsection{Threats to Validity}
%---------------------------------------------------
We have reported a favorable assessment of the proposed method. % over a diverse set of real-world benchmarks. 
Nevertheless, we have also identified five threats to the validity of our results that can further be assessed.

@@

\textit{Simulation Tool:} Only HOMER Pro was used. The inclusion of other specialized simulation tool or even a general simulation tool that uses the same mathematical model adopted by the automated verification could change the comparative.

\textit{Temperature and Solar Irradiance Data}: Information related to temperature and solar irradiance of each case study, independently of using simulation or formal verification, come from databases available online~\citep{Temperature, Irradiance}. However, considering that riverside communities do not have weather stations, the data used in our study come from the closest municipality (Manaus in all case studies), where stations collect regularly those data. Therefore, the most accurate should be the use of weather stations in each location.

@@

Our main focus was on the method to validate PV system projects, based on software model checking, which can cover the design-space better than simulation tools. 
We have considered five case studies, ranging from $253$ W to $814$ W; three state-of-art verification engines were considered (ESBMC, CBMC, and CPAchecker); and one specialized simulation tool (HOMER Pro). The experimental results from automated verification and simulation tools were compared with information collected from dwelles of the sites, where the real PV systems were deployed. Although the verification method proposed takes longer than simulation methods, it is able to present details that lead to failures in a PV system that is not a feature presented in commercial simulation tool. In particular, the proposed method was successful in finding sizing errors and indicating in details. Related to the verification engines comparative, the ESBMC with the Z3 solver executed in the incremental configuration presented the better performance (around four times faster than CBMC), used less RAM memory (less than 2 GB when compared to 9.2 GB of CBMC and 19.2 GB of CPAchecker), and all the results were sound because the PV owners and the monitoring system validated the possible flaws that the system could be presenting in the field.

@@

As future work, we will expand the number of case studies (preferentially in different countries), develop the code of a general purpose simulation tool to include in the comparative, and also consider other types of renewable energy and even hybrid ones to allow our method to verify typical rural electrification.
